---
title: "PM566 Final Project"
author: "Haoran Zhang"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---
This is my final report website. The additional data is provided in the Additional Data page and the pdf is also contained in this repository.
<br>

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}

library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)
library(readr)
library(skimr)
library(lubridate)
library(leaflet)
library(rgdal)
library(viridis)

# Initialize code chunk options
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  cache = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px",
  class.source = "code-r")
```

```{css, echo = FALSE}
.code-r { /* Code block */
  font-size: 15px;
}

.code-r-small { /* Code block */
  font-size: 10px;
}
```

<br>


# **Introduction**
Los Angeles is a very vibrant city with lots of neighborhoods. However, the crime issue is something that can not be ignored. The crime activities are frequent especially in neighborhoods near USC. In this study, we will be analyzing the arrest data in LA county from 2020 to present using the data provided by LA city database. 

# **Methods**
The arrest data comes from the LA city database. It contains detailed information of each arrest record starting from 2020.<br>
Report ID: ID for the arrest <br>
Report Type: BOOKING = Person is booked at a detention facility RFC = Person is cited and Released From Custody (RFC)<br>
Arrest time: Correcting to minutes <br>
Area ID/Name: The LAPD has 21 Community Police Stations referred to as Geographic Areas within the department. <br>
Age: in years <br>
Sex: M=Male, F=Female <br>
Charge/Charge Description: Reasons of arrest <br>
Lat/Lon/Location: Locations where the arrest occurred <br>
The unemployment data set was acquired from the Fred Economic database website, which contains the monthly unemployment rate collected from Current Population Survey (CPS). The Los Angeles council district area coefficients was captured from LA city website with further process using QGIS by inputting shape parameters and arrest counts as fields.

# **Main Question**
How did arrest counts change during the COVID-19 era and if there is any association between arrest counts and unemployment rate?

# **Data Wrangling**
We first merge the arrest data of 2020 and 2021 together in order to get an overall dataframe containing the arrest data since 2020. Then we modify the date variable by adding month and year variables for calculating monthly average data. We further exclude those arrest records with unknown location(both latitude and longitude equal 0) and those that are unlikely(age<8). Also we create a time category which partitions a day into four categories with equal length. 
```{r, echo=FALSE}
# Read CSV files
arrest20 <- read.csv("DATA/Arrest_Data_2020.csv")
arrest21 <- read.csv("DATA/Arrest_Data_2021.csv")
# using rbind to combine two datasets
arrest<- rbind(arrest20,arrest21)
```

```{r, echo=FALSE}
arrest$Arrest_Date<-strptime(arrest$Arrest_Date,format="%Y/%m/%d") 
arrest$Arrest_Date<-as.Date(arrest$Arrest_Date,format="%m/%d/%Y")
arrest$year <- year(arrest$Arrest_Date)
arrest$month <- month(arrest$Arrest_Date)
arrest$day <- day(arrest$Arrest_Date)
arrest$Booking_Date<-as.Date(arrest$Booking_Date,"%m/%d/%Y")
```

```{r, echo=FALSE}
arrest20<-arrest20[!(arrest20$LAT==0 | arrest20$LAT==0 | arrest20$Age<8),]
arrest21<-arrest21[!(arrest21$LAT==0 | arrest21$LAT==0 | arrest21$Age<8),]
arrest<-arrest[!(arrest$LAT==0 | arrest$LAT==0 | arrest$Age<8),]

daily_count_area<-arrest %>% 
  group_by(Arrest_Date,Area_Name) %>% 
  dplyr::summarise(total_arrest = n())
daily_count_area<-as.data.table(daily_count_area)


```

```{r, echo=FALSE}
arrest$time_cat <- as.factor(ifelse(arrest$Time < 600, 'before_dawn',
                          ifelse(arrest$Time < 1200, 'morning', 
                          ifelse(arrest$Time < 1800, 'afternoon', "evening"))))
```

# **Preliminary Results**
```{r, echo=FALSE}
booking<-c(nrow(arrest[arrest$Report_Type == "BOOKING" & arrest$time_cat=="morning",]),nrow(arrest[arrest$Report_Type == "BOOKING" & arrest$time_cat=="afternoon",]),nrow(arrest[arrest$Report_Type == "BOOKING" & arrest$time_cat=="evening",]),nrow(arrest[arrest$Report_Type == "BOOKING" & arrest$time_cat=="before_dawn",]))
RFC<-c(nrow(arrest[arrest$Report_Type == "RFC" & arrest$time_cat=="morning",]),nrow(arrest[arrest$Report_Type == "RFC" & arrest$time_cat=="afternoon",]),nrow(arrest[arrest$Report_Type == "RFC" & arrest$time_cat=="evening",]),nrow(arrest[arrest$Report_Type == "RFC" & arrest$time_cat=="before_dawn",]))
time_cat<-c("morning","afternoon","evening","before_dawn")
type_time<-data.frame(time_cat,booking,RFC)
```

```{r, echo=FALSE}
Arr <- plot_ly(y = ~total_arrest, type = "box", data = daily_count_area, color = ~Area_Name) %>%
  layout(title = "Arrest by Area",
         xaxis = list(title = "Area",
                      zeroline = FALSE),
         yaxis = list(title = "Arrest",
                      zeroline = FALSE))

```
 

### Choropleth Map of LA County Districts by Arrest Counts
The LAPD has 21 Community Police Stations referred to as Geographic Areas within the department. By inputting the geographic parameters of council districts data into QGIS and get a shapefile of LA districts, we are able to construct a choropleth map showing the daily average arrest of each district.
```{r, message=FALSE, echo=FALSE, warning=FALSE}
# Read this shape file with the rgdal library. 
LAPD <- readOGR("C:/Users/Zhang/PM-566-my-final-project/DATA", "LAPD_Divisions", verbose = FALSE)

# Create color palette 
col.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=LAPD@data$POP2005)
 
# Prepare the text for tooltips:
mytext <- paste(
    "Region ID: ", LAPD@data$PREC,"<br/>",
    "Region Name: ", LAPD@data$APREC,"<br/>",
    "Average Arrest Count: ", round(LAPD@data$Arrest, 3), 
    sep="") %>%
  lapply(htmltools::HTML)
 
# Final Map
leaflet(LAPD) %>% 
  addTiles()  %>% 
  setView(lat=34.05, lng=-118.44 , zoom=9.4) %>%
  addPolygons( 
    fillColor = ~col.pal(Arrest), 
    stroke=TRUE, 
    fillOpacity = 0.8, 
    color="white", 
    weight=0.3,
    label = mytext,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend( pal=col.pal, values=~Arrest, opacity=1, title = "Mean Arrest", position = "bottomleft" )
```
In this choropleth map, green means the district has relative arrest count and red means relative high arrest count. From the map we can see that 77th Street district has the highest averge arrest count, followed by Pacific and Central Districts.



```{r, echo=FALSE}
daily_avg<-plyr::count(arrest, var='Arrest_Date')
```

### Line chart of Daily Arrest
If we build a line chart of arrest counts along the time, we are able to know the variation of arrest counts along the time starting from 2020. 
```{r, echo=FALSE}
ggfig <- ggplot(daily_avg, aes(x=Arrest_Date, y=freq) ) +
         geom_line() + geom_smooth()+ggtitle("Daily Arrest Lineplot") + xlab("Date") + ylab("Arrest Count")
ggplotly(ggfig)
```
The arrest count decreased drastically since March 2020, while the overall arrest count remained flat until slowly increase in Janurary 2021.  However, there was an anormaly, the sudden burst in late May and early June 2020(May 30 - Jun 3). We claim that this was due to the BLM crime wave after the death of George Floyd on May 25, 2020 and we will show some evidence next.

### Boxplot of Arrest by Race
```{r, echo=FALSE}
daily_count_race<-arrest %>% 
  group_by(Arrest_Date,Descent_Code) %>% 
  dplyr::summarise(total_arrest = n()
            )
daily_count_race<-as.data.table(daily_count_race)

Arr_Race <- plot_ly(y = ~total_arrest, type = "box", data = daily_count_race, color = ~Descent_Code) %>%
  layout(title = "Arrest by Race",
         xaxis = list(title = "Race", xaxis = list(title = 'X Axis Title'),
                      zeroline = FALSE),
         yaxis = list(title = "Daily Arrest",
                      zeroline = FALSE))
Arr_Race
```
The figure depicts the boxplot of arrest counts by races. In this graph, we can observe that Hispanic, Black, and White appears to have the three highest arrest records. We will pick these three races as our sample and prove our claim that the sudden increase of arrest records was caused by the BLM crime rate.

```{r, echo=FALSE}
arrest_BHW<-arrest[arrest$Descent_Code=="B"|arrest$Descent_Code=="H"|arrest$Descent_Code=="W",]
tb_BHW<-dcast(setDT(arrest[arrest$Descent_Code=="B" | arrest$Descent_Code=="H"| arrest$Descent_Code=="W",]), Arrest_Date~Descent_Code, value.var = "Arrest_Date",length)[]


```
### Daily Arrest Lineplot of White, Hispanic and Black
Now we observe the pattern of arrest counts of Black, Hispanic and White.
```{r, echo=FALSE}
plot_ly(tb_BHW, x = ~Arrest_Date, y = ~B, name = 'Black', type = 'scatter', mode = 'lines') %>% 
  add_trace(y = ~H, name = 'Hispanic', type = 'scatter', mode = 'lines') %>% 
  add_trace(y = ~W, name = 'White', type = 'scatter', mode = 'lines') %>%
  layout(title = "Daily Arrest Lineplot of Black, Hispanic and White",
         xaxis = list(title = "Date",
                      zeroline = FALSE),
         yaxis = list(title = "Arrest Count",
                      zeroline = FALSE))
```
This line chart depicts the arrest count of Black, Hispanic and White. If we focus on the time around late May and early June 2020, we can see that the arrest counts of Black and Hispanic increased suddenly while the arrest counts of White continued to decrease. This provides strong evidence between the burst around late May and the BLM crime wave.

# **Unemployment Rate Data**
```{r, echo=FALSE}
# Daily arrests record table

arrest_by_date<-arrest %>% 
  group_by(Arrest_Date) %>% 
  dplyr::summarise(new_arrest = n())
arrest_by_date<-as.data.frame(arrest_by_date)
arrest_by_date<-transform(arrest_by_date,cummulation_arrest=cumsum(new_arrest))


# Compute monthly average arrest record
arrest_by_date$month<-month(arrest_by_date$Arrest_Date)
arrest_by_date$year<-year(arrest_by_date$Arrest_Date)

month_avg <- arrest_by_date %>%
  group_by(month, year) %>%
  dplyr::summarise(Mean_arrest = mean(new_arrest) )
month_avg<-month_avg %>%
  mutate(date = make_date(year, month))

unemp <- read.csv("DATA/CALOSA7URN.csv")
unemp$DATE<-strptime(unemp$DATE,format="%Y/%m/%d")
unemp$DATE<-as.Date(unemp$DATE,format="%m/%d/%Y")
unemp$year <- year(unemp$DATE)
unemp$month <- month(unemp$DATE)



unemp20_21<-unemp %>% filter(
  year>=2020
  )

unemp20_21 = subset(unemp20_21, select = -c(month,year) )
unemp_arr<-merge(x=month_avg,y=unemp20_21, by.x="date",by.y="DATE",all.x=FALSE,all.y=TRUE)
unemp_arr$year<-as.character(unemp_arr$year)
```

### Plot of unemployment rate and mean arrest
```{r, echo=FALSE}
unemp_arr[] %>%
  ggplot(mapping = aes(x = Unemployment_Rate, y = Mean_arrest, color = year, linetype = year)) +
  geom_point() + 
  geom_smooth(method = lm, se = FALSE)+
  xlab("Unemployment Rate") + ylab("Monthly Average Arrest Count")+
  ggtitle("Relation between Unemployment Rate and Arrest COunt")
```

Beyond expectation, the regression line shows a negative association between unemployment rate and arrest counts. First, the unemployment rate was low while the arrest count was high before the pandemic, which can be seen on the left up corner of the graph. Then during the lockdown, the unemployment rate grew very high, but due to the lockdown and the lack of population flow after the lock down, the arrest count decreased, such plots can be seen on the right down corner of the graph. In all, the lock down and lack of population flow lead to a negative association. 

# **Conclusion**
Among the overall arrest record from 2020 to present, most arrests were recorded in the afternoon and evening, those in the afternoon were more likely to be released from custody, while those in the evening and before dawn are more likely to be booked into retention. Number of males that is arrested is much higher than the number of females. By age, males between 20-40 years old contribute the most record. 77th Street District has the largest arrest counts follow by Pacific and Central. By races, Hispanic, Black, and White had the highest number of arrest record, maybe because of their large population base. The overall arrest count was decreasing at the start of the pandemic, probably caused by the lockdown, with sudden bursts of arrests due to the BLM crime wave around June 2020. If we combine the arrest record with the unemployment rate, we can see that the arrest record decreased as the unemployment rate increased surprisingly. Maybe this is because of the lockdown not only caused unemployment, but also caused a large decrease of population flow. Few people on street maybe associate with less crime which finally lead to less arrest records.