---
title: "PM566 Final Addition"
author: "Haoran Zhang"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

<br>

This is the additional data of my PM566 Final Project. The formal report is in the homepage of this website.

(Your output should look something like this)

<br>

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)
library(readr)
library(skimr)
library(lubridate)
library(leaflet)
library(rgdal)
library(viridis)


# Initialize code chunk options
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  cache = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px",
  class.source = "code-r")
```

```{css, echo = FALSE}
.code-r { /* Code block */
  font-size: 15px;
}

.code-r-small { /* Code block */
  font-size: 10px;
}
```

<br>


## Introduction
Los Angeles is a very vibrant city with lots of neighborhoods. However, the crime issue is something that can not be ignored. The crime activities are frequent especially in neighborhoods near USC. In this study, we will be analyzing the arrest data in LA county from 2020 to present using the data provided by LA city database. 

## Methods
The arrest data comes from the LA city database. It contains detailed information of each arrest record starting from 2020.<br>
Report ID: ID for the arrest <br>
Report Type: BOOKING = Person is booked at a detention facility RFC = Person is cited and Released From Custody (RFC)<br>
Arrest time: Correcting to minutes <br>
Area ID/Name: The LAPD has 21 Community Police Stations referred to as Geographic Areas within the department. <br>
Age: in years <br>
Sex: M=Male, F=Female <br>
Charge/Charge Description: Reasons of arrest <br>
Lat/Lon/Location: Locations where the arrest occurred <br>

The unemployment dataset also comes from LA city database which contains the unemployment rate from 2000.
## Main Question
How did arrest counts change during the COVID-19 era and if there is any association bewtween arrest counts and unemployment rate?

## Data Wrangling
Combining and cleaning data
```{r }
# Read CSV files
arrest20 <- read.csv("DATA/Arrest_Data_2020.csv")
arrest21 <- read.csv("DATA/Arrest_Data_2021.csv")
# using rbind to combine two datasets
arrest<- rbind(arrest20,arrest21)
```
Modify date variables & adding month and year variables for further use 
```{r}
arrest$Arrest_Date<-strptime(arrest$Arrest_Date,format="%Y/%m/%d") 
arrest$Arrest_Date<-as.Date(arrest$Arrest_Date,format="%m/%d/%Y")
arrest$year <- year(arrest$Arrest_Date)
arrest$month <- month(arrest$Arrest_Date)
arrest$day <- day(arrest$Arrest_Date)
arrest$Booking_Date<-as.Date(arrest$Booking_Date,"%m/%d/%Y")
```
Exclude arrests with unknown locations & those age<8
```{r}
arrest20<-arrest20[!(arrest20$LAT==0 | arrest20$LAT==0 | arrest20$Age<8),]
arrest21<-arrest21[!(arrest21$LAT==0 | arrest21$LAT==0 | arrest21$Age<8),]
arrest<-arrest[!(arrest$LAT==0 | arrest$LAT==0 | arrest$Age<8),]

daily_count_area<-arrest %>% 
  group_by(Arrest_Date,Area_Name) %>% 
  dplyr::summarise(total_arrest = n())
daily_count_area<-as.data.table(daily_count_area)


```
Time Cat & Age Cat
```{r}
arrest$time_cat <- as.factor(ifelse(arrest$Time < 600, 'before_dawn',
                          ifelse(arrest$Time < 1200, 'morning', 
                          ifelse(arrest$Time < 1800, 'afternoon', "evening"))))
arrest$age_cat <- as.factor(ifelse(arrest$Age < 20, '<20',
                          ifelse(arrest$Age < 30, '20-30',
                          ifelse(arrest$Age < 40, '30-40',
                          ifelse(arrest$Age < 50, '40-50',
                                 ">50")))))
```
## Plotting
```{r}
booking<-c(nrow(arrest[arrest$Report_Type == "BOOKING" & arrest$time_cat=="morning",]),nrow(arrest[arrest$Report_Type == "BOOKING" & arrest$time_cat=="afternoon",]),nrow(arrest[arrest$Report_Type == "BOOKING" & arrest$time_cat=="evening",]),nrow(arrest[arrest$Report_Type == "BOOKING" & arrest$time_cat=="before_dawn",]))
RFC<-c(nrow(arrest[arrest$Report_Type == "RFC" & arrest$time_cat=="morning",]),nrow(arrest[arrest$Report_Type == "RFC" & arrest$time_cat=="afternoon",]),nrow(arrest[arrest$Report_Type == "RFC" & arrest$time_cat=="evening",]),nrow(arrest[arrest$Report_Type == "RFC" & arrest$time_cat=="before_dawn",]))
time_cat<-c("morning","afternoon","evening","before_dawn")
type_time<-data.frame(time_cat,booking,RFC)

type_time_fig <- plot_ly(type_time, x = ~time_cat, y = ~booking, type = 'bar', name = 'BOOKING')
type_time_fig <- type_time_fig %>% add_trace(y = ~RFC, name = 'RFC')
type_time_fig <- type_time_fig %>% layout(yaxis = list(title = 'Count'), barmode = 'Arrest Type') %>%
   layout(title = "Types of Arrest by Time Category",
         xaxis = list(title = "Time Category",
                      zeroline = FALSE),
         yaxis = list(title = "Count",
                      zeroline = FALSE))

type_time_fig
```
This graph depicts a  barchart of time of arrest by type of arrest. Time of arrest is categorized by partitioning a whole day into 4 categories with equal length starting from midnight. There are two arrest types, booking means the person is booked at a detention facility and RFC means the person is released from custody. From this figure, we can see that despite the number of arrest is the least in the before dawn time category, the ratio of booking comparing to RFC is the largest, indicating that persons are more likely to be booked at a detention facility for arrests occur at night, while persons have higher probability to be released if arrests occur in the day. 


### Basic summary of the data
```{r}
arrest_by_area<-arrest %>% 
  group_by(Area_Name,Arrest_Date) %>% 
  dplyr::summarise(new_arrest = n())
arrest_by_area<-as.data.frame(arrest_by_area)
arrest_by_area<-transform(arrest_by_area,cummulation_arrest=cumsum(new_arrest))

dist_tb <- arrest_by_area %>%
  group_by(Area_Name) %>%
  dplyr::summarise(Mean_arrest = mean(new_arrest),
                   Max_arrest = max(new_arrest),
                   Min_arrest = min(new_arrest),
                   std_arrest = sd(new_arrest),
                   )
knitr::kable(dist_tb)
```


```{r}
Arr <- plot_ly(y = ~total_arrest, type = "box", data = daily_count_area, color = ~Area_Name) %>%
  layout(title = "Arrest by Area",
         xaxis = list(title = "Area",
                      zeroline = FALSE),
         yaxis = list(title = "Arrest",
                      zeroline = FALSE))

Arr
```
The LAPD has 21 Community Police Stations referred to as Geographic Areas within the department. This figure shows the boxplot of daily arrest 
group by the 21 police stations. From the figure we can see that 77th street has the most number of reported cases followed by Central division. 
```{r}
# Read this shape file with the rgdal library. 
LAPD <- readOGR("C:/Users/Zhang/PM-566-my-final-project/DATA", "LAPD_Divisions"
)

# Create color palette 
col.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=LAPD@data$POP2005)
 
# Prepare the text for tooltips:
mytext <- paste(
    "Region ID: ", LAPD@data$PREC,"<br/>",
    "Region Name: ", LAPD@data$APREC,"<br/>",
    "Average Arrest Count: ", round(LAPD@data$Arrest, 3), 
    sep="") %>%
  lapply(htmltools::HTML)
 
# Final Map
leaflet(LAPD) %>% 
  addTiles()  %>% 
  setView(lat=34.05, lng=-118.44 , zoom=9.4) %>%
  addPolygons( 
    fillColor = ~col.pal(Arrest), 
    stroke=TRUE, 
    fillOpacity = 0.8, 
    color="white", 
    weight=0.3,
    label = mytext,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend( pal=col.pal, values=~Arrest, opacity=1, title = "Mean Arrest", position = "bottomleft" )
```

The map of 21 LAPD regions which I'll make a choropleth map.

```{r}
library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Copy data.frames to database
dbWriteTable(con, "arrest", arrest)
dbWriteTable(con, "daily_count_area", daily_count_area)
```

```{r}
la_avg<-dbGetQuery(con, "select Area_Name, avg(total_arrest) as avg_count from daily_count_area group by Area_Name")
daily_avg<-dbGetQuery(con, "select Arrest_Date, avg(total_arrest) as avg_count from daily_count_area group by Arrest_Date")
dbWriteTable(con, "la_avg", la_avg)
dbWriteTable(con, "daily_avg", daily_avg)
```
```{r}
dbDisconnect(con)
```

```{r}
daily_avg<-plyr::count(arrest, var='Arrest_Date')

```



## Preliminary Results
```{r}
plot_ly(daily_avg, x = ~Arrest_Date, y = ~freq, type = 'scatter', mode = 'lines') %>%
  layout(title = "Daily Arrest Lineplot",
         xaxis = list(title = "Date",
                      zeroline = FALSE),
         yaxis = list(title = "Arrest Count",
                      zeroline = FALSE))
```
```{r}
p <- qplot(Arrest_Date, freq, data=daily_avg)
p <- p + geom_smooth(method = "glm", formula = y~x)

fig <- ggplotly(p)

fig
ggfig <- ggplot(daily_avg, aes(x=Arrest_Date, y=freq) ) +
         geom_line() + geom_smooth()+ggtitle("Smooth the lineplot") + xlab("2") + ylab("3")
ggplotly(ggfig)
```

This graph shows the scatter plot of arrest counts along timeline starting from 2020. Noticing the sudden increase of arrests around late June 2020. <br>
Claim: This was caused by the BLM effect.

```{r}
daily_count_race<-arrest %>% 
  group_by(Arrest_Date,Descent_Code) %>% 
  dplyr::summarise(total_arrest = n()
            )
daily_count_race<-as.data.table(daily_count_race)

Arr_Race <- plot_ly(y = ~total_arrest, type = "box", data = daily_count_race, color = ~Descent_Code) %>%
  layout(title = "Arrest by Race",
         xaxis = list(title = "Race", xaxis = list(title = 'X Axis Title'),
                      zeroline = FALSE),
         yaxis = list(title = "Daily Arrest",
                      zeroline = FALSE))
Arr_Race
```

The figure depicts the boxplot of arrest counts by races. In this graph, we can observe that Hispanic, Black and White appears to have the three highest arrest record. We will pick these three races as our sample.

```{r}
arrest_BHW<-arrest[arrest$Descent_Code=="B"|arrest$Descent_Code=="H"|arrest$Descent_Code=="W",]
tb_BHW<-dcast(setDT(arrest[arrest$Descent_Code=="B" | arrest$Descent_Code=="H"| arrest$Descent_Code=="W",]), Arrest_Date~Descent_Code, value.var = "Arrest_Date",length)[]

#tb_BHW_my<-tb_BHW
#tb_BHW_my$month<-month(tb_BHW$Arrest_Date)
#tb_BHW_my$year<-year(tb_BHW$Arrest_Date)

#dt_BHW_my <- tb_BHW_my %>%
  #group_by(month, year) %>%
 # summarise(Mean_arrest_Black = mean(B),
            #Mean_arrest_Hispanic = mean(H),
            #Mean_arrest_White = mean(W)
            #)
#dt_BHW_my<-dt_BHW_my %>%
  #mutate(date = make_date(year, month))
#dt_BHW_my<-as.data.table(dt_BHW_my)
#dt_BHW_my = subset(dt_BHW_my,select = -c(date))

plot_ly(tb_BHW, x = ~Arrest_Date, y = ~B, name = 'Black', type = 'scatter', mode = 'lines') %>% 
  add_trace(y = ~H, name = 'Hispanic', type = 'scatter', mode = 'lines') %>% 
  add_trace(y = ~W, name = 'White', type = 'scatter', mode = 'lines') %>%
  layout(title = "Daily Arrest Lineplot of Black, Hispanic and White",
         xaxis = list(title = "Date",
                      zeroline = FALSE),
         yaxis = list(title = "Arrest Count",
                      zeroline = FALSE))
```
This figure shows the lineplot of arrest counts of Hispanic, Black and White. We can see that the arrest records of Hispanic and Black did increase significantly around June 2020 while that of the Whites decreased. This supports our claim that the sudden increase of arrest counts was due to the BLM crime wave. 

```{r}
tb<-dcast(setDT(arrest), Arrest_Date~Descent_Code, value.var = "Arrest_Date",length)[]

tb_my<-tb
tb_my$month<-month(tb$Arrest_Date)
tb_my$year<-year(tb$Arrest_Date)

arrest %>% 
  group_by_if(is.numeric %>% Negate) %>%
  summarize_all(sum)

dt_my <- tb_my %>%
  group_by(month, year) %>%
  summarise(Mean_arrest_Black = mean(B),
            Mean_arrest_Hispanic = mean(H),
            Mean_arrest_White = mean(W)
            )
dt_my<-dt_my %>%
  mutate(date = make_date(year, month))
dt_my<-as.data.table(dt_my)
dt_my = subset(dt_my,select = -c(date))

colors <- c("Black" = "blue", "Hispanic" = "chocolate1", "White" = "green")
ggplot(tb, aes(x=Arrest_Date)) +
    geom_smooth(aes(y = B, color = "Black")) +
    geom_smooth(aes(y = H, color = "Hispanic")) +
    geom_smooth(aes(y = W, color = "White")) +
  labs(title = "Daily Arrest for BHW", x = "Date",
         y = "Count",
         color = "Legend") +
    scale_color_manual(values = colors)
```
In this figure which we smooth the lineplot, the arrest counts decrease in the first several months of the pandemic but started to increase starting from this year.

## Unemployment
Wrangling data
```{r}
# Daily arrests record table

arrest_by_date<-arrest %>% 
  group_by(Arrest_Date) %>% 
  dplyr::summarise(new_arrest = n())
arrest_by_date<-as.data.frame(arrest_by_date)
arrest_by_date<-transform(arrest_by_date,cummulation_arrest=cumsum(new_arrest))


# Compute monthly average arrest record
arrest_by_date$month<-month(arrest_by_date$Arrest_Date)
arrest_by_date$year<-year(arrest_by_date$Arrest_Date)

month_avg <- arrest_by_date %>%
  group_by(month, year) %>%
  dplyr::summarise(Mean_arrest = mean(new_arrest) )
month_avg<-month_avg %>%
  mutate(date = make_date(year, month))

unemp <- read.csv("DATA/CALOSA7URN.csv")
unemp$DATE<-strptime(unemp$DATE,format="%Y/%m/%d")
unemp$DATE<-as.Date(unemp$DATE,format="%m/%d/%Y")
unemp$year <- year(unemp$DATE)
unemp$month <- month(unemp$DATE)



unemp20_21<-unemp %>% filter(
  year>=2020
  )

unemp20_21 = subset(unemp20_21, select = -c(month,year) )
unemp_arr<-merge(x=month_avg,y=unemp20_21, by.x="date",by.y="DATE",all.x=FALSE,all.y=TRUE)
unemp_arr$year<-as.character(unemp_arr$year)
```
### Line Plot of Unemployment Rate
```{r}
plot_ly(unemp, x = ~DATE, y = ~Unemployment_Rate, type = 'scatter', mode = 'lines') %>%
  layout(title = "Unemployment Rate Lineplot",
         xaxis = list(title = "Date",
                      zeroline = FALSE),
         yaxis = list(title = "Unemployment Rate",
                      zeroline = FALSE))
```
From the plot above, we can see that the unemployment rate fluctuated since 1990 but increased suddenly in early 2020. We will analyze if there is some association between the arrest count and unemployment rate as intuitively, high unemployment rate could cause an increase of crime rate which would result in rise of arrest count. 
 
## Plot of unemployment rate and mean arrest
```{r}
unemp_arr[] %>%
  ggplot(mapping = aes(x = Unemployment_Rate, y = Mean_arrest, color = year, linetype = year)) +
  geom_point() + 
  geom_smooth(method = lm, se = FALSE)+
  xlab("Unemployment Rate") + ylab("Monthly Average Arrest Count")+
  ggtitle("Relation between Unemployment Rate and Arrest COunt")
```


Beyond expectation, the regression line shows a negative association between unemployment rate and arrest counts. First, the unemployment rate was low while the arrest count was high before the pandemic, which can be seen on the left up corner of the graph. Then during the lockdown, the unemployment rate grew very high, but due to the lockdown and the lack of population flow after the lock down, the arrest count decreased, such plots can be seen on the right down corner of the graph. In all, the lock down and lack of population flow lead to a negative association. 

## Conclusion
Among the overall arrest record from 2020 to present, most arrests were recorded in the afternoon and evening, those in the afternoon were more likely to be released from custody, while those in the evening and before dawn are more likely to be booked into retention. Number of males that is arrested is much higher than the number of females. By age, People between 20-40 years old contributed the most record. By races, Hispanic, Black and White had the highest number of arrest record, maybe because of their large population base. The generalized location where each race got caught also differed, which makes sense as Koreans are more concentrated in K town. The overall arrest count was decreasing at the start of the pandemic, probably caused by the lockdown, with sudden bursts of arrests due to the BLM crime wave around June 2020. If we combine the arrest record with the unemployment rate, we can see that the arrest record decreased as the unemployment rate increased suprisingly. Maybe this is because of the lockdown not only caused unemployment, but also caused a large decrease of population flow. Few people on street maybe associate with less crime which finally lead to less arrest records.








<br>
<br>